Using Dynamic Dataflow Tracking to Reduce Overheads for \\ Parallel Run-Time Monitoring

### 2014-05-23 filter.tex
Edits on Sections 1-3.

\section{Reducing Monitoring Overheads}
\label{sec:filter}

The instruction-grained monitoring techniques we consider forward all
instructions of certain types (e.g., load instructions) to the monitoring core.
However, not all of these forwarded instructions have relevant monitoring to be
done. For example, for array bounds check, all load and store instructions are
forwarded. However, only loads and stores associated with array pointers have
relevant checking to be done. Thus, we could greatly reduce the amount of
monitoring done by only forwarding instructions that correspond to data that
correspond to array pointers and have been marked with base and bound metadata.
More generally, we can reduce monitoring overheads by filtering out certain
events that correspond to empty (i.e., uninitialized) metadata.

@@

In this section, we describe how a DIFT-like hardware engine can be used to track
whether metadata has been initialized. We use this information to filter out
monitoring operations for uninitialized metadata. This engine sits between the
main core and the monitoring core (see Figure~\ref{fig:filter.overview}).

\subsection{Initialized Metadata}

% Detailed architecture of dataflow engine
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/dataflow_architecture.pdf}
    \vspace{-0.3in}
    \caption{Hardware architecture of dataflow engine.}
    \label{fig:filter.architecture} 
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Figure~\ref{fig:filter.architecture} shows a block diagram of the hardware for keeping track of
initialized metadata. When metadata is initialized, the monitor sends a signal
to the DIFT engine to mark a corresponding initialization flag. Similarly, when
the monitor propagates metadata, it also sends a signal to initialize the
metadata that has been written to. When a new event arrives at the DIFT engine,
it will read in up to two initialization flags. A table is used to configure
which initialization flags should be read in depending on the instruction type of
the event and information from the monitoring event. A pair of simple ALUs is
used to allow calculating memory addresses based on information from the
monitoring event. Typically, the initialization flags read correspond to the
source operands of the monitored instruction.

Given this set of initialization flags, a Filter Decision Table decides, based
on instruction type, whether an event can be filtered out. This table is
configured based on the specific semantics of a monitoring technique.
Typically, if all initialization flags are marked as uninitialized, then the
event can be filtered out. For example, for array bounds check, if the source
register's metadata is marked as uninitialized, then this means that the
register does not correspond to an array pointer. Thus, we can filter out
performing monitoring for this event.

\subsection{Propagating Initialization Flags}

If the source operands are marked as uninitialized, then typically we can
filter out this event. However, in these cases, the monitor also typically
resets the destination operand's metadata as cleared or uninitialized.
Thus, when we filter out one of these events, we need to also propagate the
uninitialized flag to the destination. The Filter Decision Table includes this
information about whether the destination operand metadata should be marked as
uninitialized.

>>>>

\section{Filtering Monitoring Events}
\label{sec:filter}

The instruction-grained monitoring techniques we consider forward all
instructions of certain types (e.g., load instructions) to the monitoring core.
However, not all of these forwarded instructions have relevant monitoring to be
done. For example, for array bounds check, all load and store instructions are
forwarded. However, only loads and stores associated with array pointers have
relevant monitoring operations to be done. Thus, we can greatly reduce the amount of
monitoring done by only forwarding instructions to the monitoring core that 
correspond to array pointers. These accesses can be identified because their
corresponding metadata has been set with base and bound information.
More generally, we can reduce monitoring overheads by filtering out 
events that correspond to clean or uninitialized metadata.

@@

In this section, we describe how we can use a dataflow tracking engine to
efficiently track clean metadata. This engine sits between the main core and
the monitoring core (see Figure~\ref{fig:filter.overview}) in order to filter
out monitoring operations for clean metadata.

\subsection{Tracking Clean Metadata}

% Detailed architecture of dataflow engine
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/dataflow_architecture.pdf}
    \vspace{-0.3in}
    \caption{Hardware architecture for dataflow engine.}
    \label{fig:filter.architecture} 
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Figure~\ref{fig:filter.architecture} shows a block diagram of the dataflow
tracking engine for
keeping track of clean metadata. Initially, all metadata is marked as clean.
When the monitoring core writes a non-null metadata value, it also sends a
signal to the dataflow engine to mark the metadata as dirty.

When a monitoring event arrives at the dataflow engine,
it reads in up to two clean flags. A table is used to configure
which clean flags should be read in depending on the instruction type of
the event and information from the monitoring event. A pair of simple ALUs are
used to allow calculating memory addresses based on information from the
monitoring event. Typically, the clean flags read correspond to the
source operands of the monitored instruction.

Given this set of clean flags, a Filter Decision Table decides, based
on instruction type, whether an event can be filtered out. This table is
configured based on the specific semantics of a monitoring technique.
Typically, if all clean flags are marked as clean, then the
event can be filtered out. For example, for array bounds check, if the source
register's metadata is marked as clean, then this means that the
register does not correspond to an array pointer. Thus, we can filter out the
monitoring operation for this event. 
It is important that we also propagate the clean information to the destination
of the instruction.  The Filter Decision Table also includes a bit indicating
whether the destination operand metadata should be marked as uninitialized.

### 2014-05-23 introduction.tex
Edits on Sections 1-3.

% Programmable hardware monitors
These run-time monitors introduce performance, power, and energy overheads.
Implementing run-time monitoring in software using binary instrumentation or
other similar methods introduces high overheads. For example,
dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
slowdown \cite{lift-micro06}. Implementing monitors in hardware greatly decreases
the performance impact by performing monitoring in parallel to a program's
execution. A dedicated hardware implementation of DIFT reduces performance
overheads to just 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
the programmability of software, and cannot be updated or changed.

@@

In this paper, we present an architecture that combines hardware-based dynamic
information flow tracking with a software-based monitor in order to reduce the
overheads of monitoring. By using the DIFT engine to filter out events that do
not correspond to valid metadata, we can greatly reduce the amount of
monitoring that must be handled by the monitor. For example, for an array
bounds check, we are able to filter out instruction that do not correspond to
pointer data. Since the monitor is implemented in software, the system allows a
large range of monitors to be implemented and even updated or changed at
run-time.

With this filtering, we see overheads of XX\% - XX\% on average for
uninitialized memory check, array bounds check, and multi-bit dynamic
information flow tracking. Although full monitoring coverage is always desired,
if these overheads are considered too high by the designer or user then
monitoring must be disabled.
% % Current schemes are all-or-nothing: suffer full performance impact for full
% % coverage.
% Traditionally, these overheads have been an all-or-nothing cost. If the
% designer finds that the overheads exceed their desired budget, then monitoring
% cannot be enabled.  In this paper, we present an architecture that enables a
% trade-off between the overheads and the coverage of a monitoring technique by
% dynamically adjusting the number of monitoring operations.  This system allows
% monitoring to still be performed even when the overhead budget is not
% sufficient for full monitoring.
% Use cases
Trading off coverage for
reduced overheads can still be very useful. For example, this adjustable
monitoring enables a level of protection even for systems where the full
monitoring overheads are too high.  This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines. 
Additionally, adjustable monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05,
chilimbi-asplos04, greathouse-cgo11}.

Thus, we develop an architecture that enables a
trade-off between the overheads and the coverage of a monitoring technique by
dynamically adjusting the number of monitoring operations. By extending the
hardware DIFT engine to track 2-bits of information, we are able to use it for
filtering empty metadata as well as tracking dropped monitoring flows in order
to prevent false positives. We introduce three policies of where dropping
decisions are made and show the trade-off space achieved between accuracy of
overhead matching and coverage achieved.

>>>>

% Programmable hardware monitors
These run-time monitors introduce performance, power, and energy overheads.
Implementing run-time monitoring in software using binary instrumentation or
other similar methods has been shown to introduce high overheads. For example,
dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
slowdown \cite{lift-micro06}. Implementing monitors in hardware greatly decreases
the performance impact by performing monitoring in parallel to a program's
execution. A dedicated hardware implementation of DIFT reduces performance
overheads to just 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
the programmability of software, and cannot be updated or changed.

@@

In this paper, we present an architecture that combines a hardware-based
dataflow tracking engine with a software-based monitor in order to reduce the
overheads of monitoring. By using the dataflow engine to track flows of
initialized metadata, the system is able to filter out events that do not
correspond to valid metadata, greatly reducing the amount of monitoring that
must be handled in software. For example, for an array
bounds check, the dataflow engine is able to filter out instruction that do not
correspond to pointer data. Thus, we are able to have the flexibility of
software-based monitors with greatly reduced overheads.

With this filtering, we see overheads of XX\% - XX\% on average for
uninitialized memory check, array bounds check, and multi-bit dynamic
information flow tracking. Although full monitoring coverage is always desired,
if these overheads are considered too high by the designer or user then
monitoring must be disabled.
% % Current schemes are all-or-nothing: suffer full performance impact for full
% % coverage.
% Traditionally, these overheads have been an all-or-nothing cost. If the
% designer finds that the overheads exceed their desired budget, then monitoring
% cannot be enabled.  In this paper, we present an architecture that enables a
% trade-off between the overheads and the coverage of a monitoring technique by
% dynamically adjusting the number of monitoring operations.  This system allows
% monitoring to still be performed even when the overhead budget is not
% sufficient for full monitoring.
% Use cases
Thus, we present an architecture that enables a trade-off between overheads and
the monitoring coverage achieved. This is achieved by dynamically dropping
monitoring operations when overheads are too high. We are able to prevent false
positives by extending the dataflow engine to also track these dropped
monitoring flows. We also present three policies on when dropping decisions are
made that cover a trade-off space between ability to match a target overhead
and the monitoring coverage achieved.

Although full monitoring coverage is always desired, trading off coverage for
reduced overheads can still be very useful. For example, this adjustable
monitoring enables a level of protection even for systems where the full
monitoring overheads are too high. This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines.
Additionally, adjustable monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05, chilimbi-asplos04,
greathouse-cgo11}.

The main contributions of this paper are as follows.
\begin{itemize}
  \item Using a dataflow tracking engine to filter out monitoring events in order to reduce overheads.
  \item Extending the dataflow tracking engine to enable monitoring with adjustable overheads
  \item Exploration of the design space and trade-offs in dropping policies for performing partial monitoring
\end{itemize}

### 2014-05-23 monitoring.tex
Edits on Sections 1-3.

Figure~\ref{fig:arch.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitor} performs a
series of \emph{monitoring operations}. The monitor operates in parallel to the
main core. These events are referred to as \emph{monitoring events}. Depending
on the type of a monitoring event, different monitoring operations are
executed. Monitoring events are buffered in a FIFO structure to decouple the
running of the main core and the monitor. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. For programmable monitors, these stalls are a major source of
overhead because the monitor may take several cycles to process a single event
from the main core. We refer to these stalls and other overheads, such as
contention for shared resources, as \emph{monitoring overheads}. If the monitor
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

There are many possible monitoring schemes that can be implemented on this type
of fine-grained parallel monitoring architecture such as memory protection
\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
\cite{hardbound-asplos08} can be implemented where monitoring is used to detect
when software attempts to read or write to a memory location outside of an
array's bounds. This can be done by associating base and bound address metadata
to an array pointer when it is allocated. On loads or stores with the array
pointer, the monitor checks that the memory address is within the base and
bound. In addition, on certain ALU instructions, this base and bound metadata
is propagated to track the creation of new array pointers.
% One example is an uninitialized memory check (UMC) where monitoring is used to
% detect when software attempts to read from a memory location that was not
% previously initialized. This can be done by forwarding each load and store
% instruction from the main core to the monitor. For every memory location, the
% monitor keeps one bit of metadata. On a store to a memory location, the monitor
% marks the corresponding metadata bit to indicate that the memory location has
% been initialized. On a load, the monitor checks that the corresponding metadata
% bit has been previously marked as initialized.

There are multiple options for implementing programmable parallel monitors. For
example, the Log-Based Architecture \cite{lba-isca08} uses processor cores in a
multi-core system as monitors. The FlexCore architecture
\cite{flexcore-micro10} instead uses an FPGA-fabric to implement the
monitor. The approach we describe in this paper applies to any of these
parallel monitors. However, for experiments, we model an FPGA-based monitor
similar to FlexCore.  The on-chip FPGA fabric is used to implement the
``Monitor'' block in Figure~\ref{fig:arch.overview} while the FIFO from the
main core and metadata cache are implemented as ASICs.

>>>>

Figure~\ref{fig:arch.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitoring core} performs a
series of \emph{monitoring operations}. The monitoring core operates in parallel to the
main core. These events are referred to as \emph{monitoring events}. Depending
on the type of a monitoring event, different monitoring operations are
executed. Monitoring events are buffered in a FIFO structure to decouple the
running of the main core and the monitoring core. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. These stalls are a major source of
overhead because the monitoring core may take several cycles to process a single event
from the main core. We refer to these stalls and other overheads, such as
contention for shared resources, as \emph{monitoring overheads}. If the monitoring core
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

There are many possible monitoring schemes that can be implemented on this type
of fine-grained parallel monitoring architecture such as memory protection
\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
\cite{hardbound-asplos08} can be implemented where monitoring is used to detect
when software attempts to read or write to a memory location outside of an
array's bounds. This can be done by associating base and bound address metadata
to an array pointer when it is allocated. On loads or stores with the array
pointer, the monitoring core checks that the memory address is within the base and
bound. In addition, on certain ALU instructions, this base and bound metadata
is propagated to track the creation of new array pointers.
% One example is an uninitialized memory check (UMC) where monitoring is used to
% detect when software attempts to read from a memory location that was not
% previously initialized. This can be done by forwarding each load and store
% instruction from the main core to the monitor. For every memory location, the
% monitor keeps one bit of metadata. On a store to a memory location, the monitor
% marks the corresponding metadata bit to indicate that the memory location has
% been initialized. On a load, the monitor checks that the corresponding metadata
% bit has been previously marked as initialized.

% There are multiple options for implementing programmable parallel monitors. For
% example, the Log-Based Architecture \cite{lba-isca08} uses processor cores in a
% multi-core system as monitors. The FlexCore architecture
% \cite{flexcore-micro10} instead uses an FPGA-fabric to implement the
% monitor. The approach we describe in this paper applies to any of these
% parallel monitors. However, for experiments, we model an FPGA-based monitor
% similar to FlexCore.  The on-chip FPGA fabric is used to implement the
% ``Monitor'' block in Figure~\ref{fig:arch.overview} while the FIFO from the
% main core and metadata cache are implemented as ASICs.

### 2014-05-23 title.tex
Edits on Sections 1-3.

Recent studies have proposed various parallel run-time monitoring techniques to
improve the reliability, security, and debugging capabilities of
computer systems. These run-time monitors introduce overheads in performance
and energy to the system. Traditionally, these overheads have been an
all-or-nothing cost; monitoring cannot be used if the overheads are considered too large.
% where if monitoring is enabled, then the program will incur, in full, the
% associated overheads. 
%That is, if the overheads are considered too large, then monitoring cannot be
%enabled. 
In this paper, we present a run-time monitoring architecture that
enables a trade-off between the coverage and the overheads.
The system allows an overhead or coverage target to be specified,
and limits the amount of monitoring performed at run-time to meet the target. 
However, simply skipping monitoring operations can lead to false
positives, signaling errors when none have occurred. Our proposed architecture
prevents false positives using a method that invalidates certain metadata
when monitoring is not performed. Dedicated hardware that performs dynamic
information flow tracking (DIFT) is used to propagate this invalidation
information in order to prevent false positives. In addition, we can use this
DIFT hardware to also track the presence of valid metadata. This allows us to
filter out events in order to reduce monitoring overheads.
% Our experiments show that the system is able to
% closely meet the target overhead and achieve significant monitoring coverage
% for reasonable overhead targets.
In our experiments, we see that the system is able to come
within 2\% of the specified execution time overhead in almost all cases.  In addition, with a
performance overhead target of 10\%, 49-90\% of monitoring coverage is achieved
on average depending on the monitoring technique.

>>>>

Recent studies have proposed various parallel run-time monitoring techniques to
improve the reliability, security, and debugging capabilities of
computer systems. These run-time monitors introduce overheads in performance
and energy to the system. Traditionally, these overheads have been an
all-or-nothing cost; monitoring cannot be used if the overheads are considered too large.
% where if monitoring is enabled, then the program will incur, in full, the
% associated overheads. 
%That is, if the overheads are considered too large, then monitoring cannot be
%enabled. 
In this paper, we present a run-time monitoring architecture that
enables a trade-off between the coverage and the overheads.
The system allows an overhead or coverage target to be specified,
and limits the amount of monitoring performed at run-time to meet the target. 
However, simply skipping monitoring operations can lead to false
positives, signaling errors when none have occurred. Our proposed architecture
prevents false positives using a method that invalidates certain metadata
when monitoring is not performed. Dedicated hardware that performs dynamic
dataflow tracking is used to propagate this invalidation
information in order to prevent false positives. In addition, we can use this
dataflow hardware to also track the presence of initialized metadata. This
allows us to filter out events on uninitialized metadata in order to reduce
monitoring overheads.
% Our experiments show that the system is able to
% closely meet the target overhead and achieve significant monitoring coverage
% for reasonable overhead targets.
In our experiments, we see that the system is able to come within 2\% of the
specified execution time overhead in almost all cases.  In addition, with a
performance overhead target of 10\%, 49-90\% of monitoring coverage is achieved
on average depending on the monitoring technique.
